---
title: "HW5 : Distribution"
author:
- Ebenezer Nkum^[enkum@miners.utep.edu]
- University of Texas at El Paso (UTEP)
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    number_sections: yes
    toc: yes
    toc_depth: 4
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
geometry: margin=1in
fontsize: 11pt
spacing: single
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsfonts}
- \usepackage{amsthm}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- "\\rhead{DS 6339-- Data Visualization}"
- \cfoot{\thepage}
- \usepackage{algorithm}
- \usepackage[noend]{algpseudocode}
- "\\DeclareUnicodeCharacter{2212}{\\ensuremath{-}}"
editor_options: 
  chunk_output_type: inline
---

<!-- QUESTION ONE: WHAT --> 
\noindent\rule{17.5cm}{0.8pt}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(plotly)
library(hrbrthemes)
library(kableExtra)
library(knitr)
library(tinytex)
library(tibble)
library(ggrepel)
library("reshape2") 
library("patchwork")
```



# Load data
```{r Load data}
# Load data 
dat <- read.csv("serialdat.csv", header = T)
```

# Data

The data contains information about gene variant transcriptions. There are three replications of the variant transcriptions and a final column where the three replications were averaged. The categorical variables included are the SUMOver (which is classes of genes - S#V# format).


# Step 1: Inspect data, assess it for completeness, good formatting, and any errors
## Inspect data
```{r inspect data, echo=FALSE}

head(dat) ; dim(dat)

```

## Remove the replication number that follows the S#V# groups
```{r , message=FALSE}
dat0 <- dat[-43,] %>% select(-X10.x.copies) %>% 
                mutate(SUMOvar = sapply(str_split(SUMOvar,"-",),'[',1))

dat0 <- dat0 %>% mutate(SUMOvar = substr(SUMOvar,1,2))


head(dat0)
```

# Visualizing distribution

# Overall distribution among each group using the average replications

```{r, fig.width=10}

par(mfrow=c(2,2))
p <- ggplot(dat0, aes(x= Replicate.1, y=Replicate.2, color = SUMOvar)) +
      geom_point()+
      theme_classic() +
      ggtitle("Relationship between Rep1\n and Rep2 by Classes of Gene")+
      theme(legend.position = "none")



p1 <- ggplot(dat0, aes(x= Replicate.1, y=Replicate.3, color = SUMOvar)) +
      geom_point()+
      theme_classic() +
      ggtitle("Relationship between Rep1\n and Rep3 by Classes of Gene")+
      theme(legend.position = "none")


p2 <- ggplot(dat0, aes(x= Replicate.2, y=Replicate.3, color = SUMOvar)) +
      geom_point()+
      theme_classic() +
      ggtitle("Relationship between Rep2\n and Rep3 by Classes of Gene") 
 
p2$labels$colour<-"Classes \n of Gene"

p+p1+p2

```
Upon examining the associations between the three transcription entities, it is apparent that there exists a significant degree of variability amongst the s2 and s3 gene classes in relation to the association between replication 1 and 2. A similar observation can be made regarding the correlation between replication 1 and replication 3. However, it is worth noting that there is very minimal variability or nearly perfect association with regards to the association between replication 2 and replication 3.













